\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
setwd("~/EDE_FinalProject/code/Raw") 
#setwd("~/EDE_FinalProject/Raw")
# Load your packages
library(ggplot2)
library(dplyr)
library(knitr)
library(tidyr)

# Set your ggplot theme
custom_theme <- theme(
  # Background theme customization
  panel.background = element_rect(fill = "white", color = "black"),  
  # Panel background color
  plot.background = element_rect(fill = "#f0f0f0", color = NA),  
  # Plot area background
  panel.grid.major = element_line(color = "gray", size = 0.5),  
  # Major grid lines
  panel.grid.minor = element_line(color = "lightgray", size = 0.25),  
  # Minor grid lines
  axis.line = element_line(color = "black"),  
  # Axis lines
  axis.text = element_text(size = 12, color = "black"),  
  # Axis tick labels
  axis.title = element_text(size = 14, face = "bold", color = "black"),  
  # Axis titles
  plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  
  # Plot title customization
  legend.position = "none",  
  # Hide legend
  strip.background = element_rect(fill = "#D3D3D3", color = "black"),  
  # Strip background for facets
  strip.text = element_text(size = 14, face = "bold")  
  # Facet label text
)



# Load your datasets
fuel <- read.csv("Fuel_data.csv")
ghgp2010 <- read.csv("ghgp_data_2010.csv")
ghgp2011 <- read.csv("ghgp_data_2011.csv")
ghgp2012 <- read.csv("ghgp_data_2012.csv")
ghgp2013 <- read.csv("ghgp_data_2013.csv")
ghgp2014 <- read.csv("ghgp_data_2014.csv")
ghgp2015 <- read.csv("ghgp_data_2015.csv")
ghgp2016 <- read.csv("ghgp_data_2016.csv")
ghgp2017 <- read.csv("ghgp_data_2017.csv")
ghgp2018 <- read.csv("ghgp_data_2018.csv")
ghgp2019 <- read.csv("ghgp_data_2019.csv")
ghgp2020 <- read.csv("ghgp_data_2020.csv")
ghgp2021 <- read.csv("ghgp_data_2021.csv")
ghgp2022 <- read.csv("ghgp_data_2022.csv")
ghgp2023 <- read.csv("ghgp_data_2023.csv")
ghgpyear <- read.csv("ghgp_data_by_year_2023.csv")
```

# Rationale and Research Questions
>With current global temperature rise accelerating towards dangerously high levels, there is a significant threat to meeting the targets established by the Paris Climate Accords. A main target for this agreement was to limit global temperature to a 2.0° increase. According to the International Energy Agency, the United States and China together accounted for 45% of global fuel combustion emissions. (IEA). With the US being such a large contributor to climate change, a major goal of climate scientists is to find ways to reduce US GHG emissions. For this analysis, we chose to look at EPA-provided emission data from 2010-2023. The main purpose of this analysis is to gain insight on which industries demonstrate the greatest potential for significant emissions reductions, and to better understand what role do facility and fuel characteristics play in driving this change over the past decade?

>Industry sector data, greenhouse gas emission data, rated heat capacity data, and geospatial data will all be used to identify relationships and areas for improvement in the US. Additionally, most of our research questions are based on the most recent data, from 2023; however, time series analysis takes into account the full range of our data. Our central research questions set to examine how certain industry sectors should be limited in their production to limit the levels of greenhouse gas (especially carbon dioxide) emissions.

\newpage

# Dataset Information -Yufan

# Data Cleaning
>We used two main datasets, as well as a couple of shapefiles of the US. For the downloaded datasets, the main issue we encountered was that some facilities included multiple industry sector types. This made it difficult to group our data, so we decided to take the most significant sector (the first sector listed for each facility) and base our findings from that. Missing data values were excluded, and year/date format was checked to ensure the time series analysis would run smoothly. For the shapefile, we excluded all areas East of 50°West longitude, as the focus was on the US. After these initial steps were taken, we had confidence in the organization of our data and were ready for analysis.

Downloaded datasets included date/time and mean discharge (cfs) columns. After the date column was properly classified as a date, unnecessary columns were removed from the data frame. Next, the date column was separated out to filter for years 1990-2021 and to filter for hurricane months (June-October). The dataset was then saved into the processed folder for each of the research locations. Missing data values were then filled using a linear interpolation and then daily averages of discharge were created in a new, clean column of daily means. After all fo these steps were taken the data was considered to be wrangled and was then ready for initial visualizations.

\newpage

# Exploratory Analysis -Yufan

## 2023 Data Overview by Industry Type (Sectors)

```{r 2023 sector summary, echo=FALSE, warning=FALSE, message=FALSE}
# Split the "Industry.Type..sectors." column by commas and unnest into separate rows
industry_summary <- ghgp2023 %>%
  separate_rows(`Industry.Type..sectors.`, sep = ",") %>%  # Correct column name
  mutate(`Industry.Type..sectors.` = trimws(`Industry.Type..sectors.`)) %>%  # Remove extra spaces
  group_by(`Industry.Type..sectors.`) %>%  # Group by the individual sectors
  summarise(Number_of_Reports = n())  # Count the number of reports for each sector

# Display the summarized data as a table
kable(industry_summary, caption = "Summary of Number of Reporters by Industry Sector")
```

## 2023 Data Overview by Greenhous Gas Type

```{r 2023 GHG summary, echo=FALSE, warning=FALSE, message=FALSE}
# Reshape the data into long format
ghgp_long <- ghgp2023 %>%
  select(
    CO2.emissions..non.biogenic., 
    Methane..CH4..emissions, 
    Nitrous.Oxide..N2O..emissions, 
    HFC.emissions, 
    PFC.emissions, 
    SF6.emissions, 
    NF3.emissions, 
    Other.Fully.Fluorinated.GHG.emissions
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Gas_Type",
    values_to = "Emissions"
  )

# View the reshaped data (first few rows)
head(ghgp_long)
```
```{r 2023 GHG plot, echo=FALSE, warning=FALSE}
# Create the bar chart with the custom theme
ggplot(ghgp_long, aes(x = Emissions, y = Gas_Type, fill = Gas_Type)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Greenhouse Gas Emissions in 2023",
    x = "Greenhouse Gas Type",
    y = "Emissions (Tons of CO2)"
  ) +
  custom_theme +  # Apply the custom theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  )

```
## 2023 Average CO2 Emissions by Industry Sector 

```{r 2023 Industry Sector plot, echo=FALSE, warning=FALSE, message=FALSE}
# Calculate average CO2 emissions by industry sector
 CO2_2023 <- ghgp2023 %>%
  separate_rows(`Industry.Type..sectors.`, sep = ",") %>%  # Correct column name
  mutate(`Industry.Type..sectors.` = trimws(`Industry.Type..sectors.`)) %>%  # Remove extra spaces
  group_by(`Industry.Type..sectors.`) %>%  # Group by the individual sectors
  summarise(Average_CO2 = mean(CO2.emissions..non.biogenic., na.rm = TRUE) / 10^6)

CO2_2023$Industry.Type..sectors. <- gsub("Fluorintaed", "Fluorinated", CO2_2023$Industry.Type..sectors.)
CO2_2023$Industry.Type..sectors. <- gsub("Import and Export", "Distribution", CO2_2023$Industry.Type..sectors.)
# Generate the bar chart
ggplot(CO2_2023, aes(x = reorder(Industry.Type..sectors., -Average_CO2), y = Average_CO2, fill = Industry.Type..sectors.)) +
  geom_bar(stat = "identity") +
  labs(title = "Average 2023 CO2 Emissions by Industry Sector",
       x = "Industry Sector",
       y = "Average CO2 Emissions (MMT)") +
  custom_theme +  # Apply the custom theme
  theme(
    axis.text.x = element_text(hjust = 1)  # Rotate x-axis labels
  )+
 coord_flip() 
```

# 2023  Total CO2 Emissions by Industry Sector and State

```{r 2023 Industry Sector by State plot, echo=FALSE, warning=FALSE, message=FALSE}

CO2_by_state_sector <- ghgp2023 %>%
  separate_rows(`Industry.Type..sectors.`, sep = ",") %>%
  mutate(`Industry.Type..sectors.` = trimws(`Industry.Type..sectors.`)) %>%
  group_by(State, `Industry.Type..sectors.`) %>% 
  summarise(Total_CO2 = sum(CO2.emissions..non.biogenic., na.rm = TRUE) / 10^6) 
  
CO2_by_state_sector$Industry.Type..sectors. <- gsub("Fluorintaed", "Fluorinated", CO2_by_state_sector$Industry.Type..sectors.)
CO2_by_state_sector$Industry.Type..sectors. <- gsub("Import and Export", "Distribution", CO2_by_state_sector$Industry.Type..sectors.)
# Generate the stacked bar chart
ggplot(CO2_by_state_sector, 
       aes(x = reorder(State, -Total_CO2), 
           y = Total_CO2, 
           fill = `Industry.Type..sectors.`)) +
  geom_bar(stat = "identity") + 
  labs(title = "Total 2023 CO2 Emissions by State and Industry Sector",
       x = "State",
       y = "Total CO2 Emissions (MMT)") +
  custom_theme + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()

```

# 2023 CO2 Emissions by State

```{r 2023 Industry Sector plot, echo=FALSE, warning=FALSE, message=FALSE}

library(sf)
# Filter out rows with Longitude greater than 100
ghgp2023 <- ghgp2023[ghgp2023$Longitude <= 60, ]

ghgp2023 <- ghgp2023[!is.na(ghgp2023$Longitude) & !is.na(ghgp2023$Latitude), ]
USnation_sf<- st_read("Raw/cb_2018_us_nation_20m.shp")
st_crs(USnation_sf)

ghgp2023_sf <- ghgp2023 %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269) 

str(ghgp2023_sf)

st_crs(ghgp2023_sf)

ggplot() +
  geom_sf(data = USnation_sf) + 
  geom_sf(data = ghgp2023_sf,
          aes(size = CO2.emissions..non.biogenic.,
              color = CO2.emissions..non.biogenic.),  # Map CO2 emissions to color 
          alpha = 0.5) + 
  labs(title = "CO2 Emissions from Gas Plants",
       size = "CO2 Emissions") +
  coord_sf(xlim = c(-180, -50)) +  # Set longitude limits
  scale_color_gradient(low = "cadetblue1", high = "grey1")+  # Define the color scale
  scale_size(range = c(1,7)) 

```
## 2023 Direct Emission Facilities by State
```{r 2023 Facilities plot, echo=FALSE, warning=FALSE, message=FALSE}
# Calculate average CO2 emissions by industry sector
units_data <- read.csv("Raw/ede_csv_unit_emissions.csv", skip = 6)
#RHC_by_state <- units_data %>% 
  #group_by(State) %>% 
  #summarise(Average_RHC = mean(Unit Maximum Rated Heat Input Capacity (mmBTU/hr), na.rm = TRUE)
            
Facilities_by_state <- units_data %>% 
  filter(`Reporting.Year` == 2023) %>% 
  group_by(State) %>% 
  summarise(FacilityCount = n_distinct(Facility.Id)) %>%
  arrange(desc(FacilityCount))

# Create the bar chart
ggplot(Facilities_by_state, aes(x = reorder(State, -FacilityCount), y = FacilityCount, fill = FacilityCount)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Number of Unique Facilities by State (2023)",
    x = "State",
    y = "Number of Facilities"
  )

```

## 2023 Average Rated Heat Capacity by State
```{r 2023 AVG Rated Heat Capacity plot, echo=FALSE, warning=FALSE, message=FALSE}
# Calculate RHC by state

AVG_RHC_by_state <- units_data %>% 
  filter(`Reporting.Year` == 2023) %>% 
  group_by(State) %>% 
  summarise(Average_RHC = mean(Unit.Maximum.Rated.Heat.Input.Capacity..mmBTU.hr., na.rm = TRUE))%>%
  arrange(desc(Average_RHC))


# Create the bar chart
ggplot(AVG_RHC_by_state, aes(x = reorder(State, -Average_RHC), y = Average_RHC, fill = Average_RHC)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Average Rated Heat Capacity by State (2023)",
    x = "State",
    y = "Rated Heat Capacity"
  )
```

## 2023 Total Rated Heat Capacity by State
```{r 2023 Total Rated Heat Capacity plot, echo=FALSE, warning=FALSE, message=FALSE}
Total_RHC_by_state <- units_data %>% 
  filter(`Reporting.Year` == 2023) %>% 
  group_by(State) %>% 
  summarise(Total_RHC = sum(Unit.Maximum.Rated.Heat.Input.Capacity..mmBTU.hr., na.rm = TRUE)) %>%
  arrange(desc(Total_RHC))
# Create the bar chart
ggplot(Total_RHC_by_state, aes(x = reorder(State, -Total_RHC), y = Total_RHC, fill = Total_RHC)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Total Rated Heat Capacity by State (2023)",
    x = "State",
    y = "Rated Heat Capacity (MMBTU/hr)"
  )
```

\newpage

# Analysis





## Question 1: Which industries demonstrate the greatest change for significant emissions reductions over the past decade? [delete later -3 in ideas]

```{r time series, echo=FALSE, warning=FALSE, message=FALSE}

```


## Question 2: What role do facility and fuel characteristics play in driving this change over the past decade?

### Facility Emission Analysis 

```{r 2023 facility, echo=FALSE, warning=FALSE}
ghgp2023 <- ghgp2023 %>%
  mutate(Emission_Range = cut(
    `Total.reported.direct.emissions`,
    breaks = c(0, 25000, 50000, 100000, 250000, 1000000, Inf),
    labels = c("0-25", "25-50", "50-100", "100-250", "250-1,000", ">1,000")
  ))

sector_summary <- ghgp2023 %>%
  separate_rows(`Industry.Type..sectors.`, sep = ",") %>%  # Split sectors
  mutate(`Industry.Type..sectors.` = trimws(`Industry.Type..sectors.`)) %>%  # Remove extra spaces
  group_by(Emission_Range, `Industry.Type..sectors.`) %>%
  summarise(
    Total_Emissions = sum(`Total.reported.direct.emissions`, na.rm = TRUE),
    .groups = "drop"
  )

range_counts <- ghgp2023 %>%
  group_by(Emission_Range) %>%
  summarise(Count = n(), .groups = "drop")

final_data <- sector_summary %>%
  left_join(range_counts, by = "Emission_Range")
```

```{r 2023 facility plot, echo=FALSE, warning=FALS}
# Plot
ggplot(final_data, aes(x = Emission_Range, y = Total_Emissions, fill = Industry.Type..sectors.)) +
  geom_bar(stat = "identity") +  # Stacked bar chart
  geom_text(data = range_counts, aes(x = Emission_Range, y = max(final_data$Total_Emissions) * 1.05, label = Count),
            inherit.aes = FALSE, size = 4) +  # Add count labels
  labs(
    title = "Facility Emission Ranges (2023)",
    x = "Emission Range of Facilities (thousand metric tons CO2e)",
    y = "2023 Emissions (million metric tons CO2e)",
    fill = "Sector"
  ) +
  theme_minimal() +
  custom_theme
```


### Fuel Type Emission Analysis





\newpage

# Summary and Conclusions


\newpage

# References
“Greenhouse Gas Emissions from Energy Data Explorer – Data Tools.” IEA, 2 Aug. 2024, www.iea.org/data-and-statistics/data-tools/greenhouse-gas-emissions-from-energy-data-explorer. 
