\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
setwd("~/EDE_FinalProject/code/Raw") 
#setwd("~/EDE_FinalProject/Raw")
# Load your packages
library(ggplot2)
library(dplyr)
library(knitr)
library(tidyr)

# Set your ggplot theme
custom_theme <- theme(
  # Background theme customization
  panel.background = element_rect(fill = "white", color = "black"),  
  # Panel background color
  plot.background = element_rect(fill = "#f0f0f0", color = NA),  
  # Plot area background
  panel.grid.major = element_line(color = "gray", size = 0.5),  
  # Major grid lines
  panel.grid.minor = element_line(color = "lightgray", size = 0.25),  
  # Minor grid lines
  axis.line = element_line(color = "black"),  
  # Axis lines
  axis.text = element_text(size = 12, color = "black"),  
  # Axis tick labels
  axis.title = element_text(size = 14, face = "bold", color = "black"),  
  # Axis titles
  plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  
  # Plot title customization
  legend.position = "none",  
  # Hide legend
  strip.background = element_rect(fill = "#D3D3D3", color = "black"),  
  # Strip background for facets
  strip.text = element_text(size = 14, face = "bold")  
  # Facet label text
)



# Load your datasets
fuel <- read.csv("Fuel_data.csv")
ghgp2010 <- read.csv("ghgp_data_2010.csv")
ghgp2011 <- read.csv("ghgp_data_2011.csv")
ghgp2012 <- read.csv("ghgp_data_2012.csv")
ghgp2013 <- read.csv("ghgp_data_2013.csv")
ghgp2014 <- read.csv("ghgp_data_2014.csv")
ghgp2015 <- read.csv("ghgp_data_2015.csv")
ghgp2016 <- read.csv("ghgp_data_2016.csv")
ghgp2017 <- read.csv("ghgp_data_2017.csv")
ghgp2018 <- read.csv("ghgp_data_2018.csv")
ghgp2019 <- read.csv("ghgp_data_2019.csv")
ghgp2020 <- read.csv("ghgp_data_2020.csv")
ghgp2021 <- read.csv("ghgp_data_2021.csv")
ghgp2022 <- read.csv("ghgp_data_2022.csv")
ghgp2023 <- read.csv("ghgp_data_2023.csv")
ghgpyear <- read.csv("ghgp_data_by_year_2023.csv")
```

# Rationale and Research Questions



\newpage

# Dataset Information



\newpage

# Exploratory Analysis [delete later -1 in ideas]

## 2023 Data Overview by Industry Type (Sectors)

```{r 2023 sector summary, echo=FALSE, warning=FALSE, message=FALSE}
# Split the "Industry.Type..sectors." column by commas and unnest into separate rows
industry_summary <- ghgp2023 %>%
  separate_rows(`Industry.Type..sectors.`, sep = ",") %>%  # Correct column name
  mutate(`Industry.Type..sectors.` = trimws(`Industry.Type..sectors.`)) %>%  # Remove extra spaces
  group_by(`Industry.Type..sectors.`) %>%  # Group by the individual sectors
  summarise(Number_of_Reports = n())  # Count the number of reports for each sector

# Display the summarized data as a table
kable(industry_summary, caption = "Summary of Number of Reporters by Industry Sector")
```

## 2023 Data Overview by Greenhous Gas Type

```{r 2023 GHG summary, echo=FALSE, warning=FALSE, message=FALSE}
# Reshape the data into long format
ghgp_long <- ghgp2023 %>%
  select(
    CO2.emissions..non.biogenic., 
    Methane..CH4..emissions, 
    Nitrous.Oxide..N2O..emissions, 
    HFC.emissions, 
    PFC.emissions, 
    SF6.emissions, 
    NF3.emissions, 
    Other.Fully.Fluorinated.GHG.emissions
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Gas_Type",
    values_to = "Emissions"
  )

# View the reshaped data (first few rows)
head(ghgp_long)
```
```{r 2023 GHG plot, echo=FALSE, warning=FALSE}
# Create the bar chart with the custom theme
ggplot(ghgp_long, aes(x = Emissions, y = Gas_Type, fill = Gas_Type)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Greenhouse Gas Emissions in 2023",
    x = "Greenhouse Gas Type",
    y = "Emissions (Tons of CO2)"
  ) +
  custom_theme +  # Apply the custom theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  )

```
## 2023 Data Overview by Greenhous Gas Type

```{r 2023 Industry Sector plot, echo=FALSE, warning=FALSE, message=FALSE}
# Calculate average CO2 emissions by industry sector
 CO2_2023 <- ghgp2023 %>%
  separate_rows(`Industry.Type..sectors.`, sep = ",") %>%  # Correct column name
  mutate(`Industry.Type..sectors.` = trimws(`Industry.Type..sectors.`)) %>%  # Remove extra spaces
  group_by(`Industry.Type..sectors.`) %>%  # Group by the individual sectors
  summarise(Average_CO2 = mean(CO2.emissions..non.biogenic., na.rm = TRUE) / 10^6)

# Generate the bar chart
ggplot(CO2_2023, aes(x = reorder(Industry.Type..sectors., -Average_CO2), y = Average_CO2, fill = Industry.Type..sectors.)) +
  geom_bar(stat = "identity") +
  labs(title = "Average 2023 CO2 Emissions by Industry Sector",
       x = "Industry Sector",
       y = "Average CO2 Emissions (MMT)") +
  custom_theme +  # Apply the custom theme
  theme(
    axis.text.x = element_text(hjust = 1)  # Rotate x-axis labels
  )+
 coord_flip() 


```
```{r 2023 Industry Sector plot, echo=FALSE, warning=FALSE, message=FALSE}
# Calculate average CO2 emissions by industry sector
RHC <- read.csv("ede_csv_unit_emissions.csv")
CO2_2023 <- RHC %>% 
  group_by(State) %>% 
  summarise(Average_RHC = mean(Unit Maximum Rated Heat Input Capacity (mmBTU/hr), na.rm = TRUE)


```

\newpage

# Analysis





## Question 1: Which industries demonstrate the greatest change for significant emissions reductions over the past decade? [delete later -3 in ideas]

```{r time series, echo=FALSE, warning=FALSE, message=FALSE}

```


## Question 2: What role do facility and fuel characteristics play in driving this change over the past decade?

### Facility Emission Analysis [delete later -5 in ideas]

```{r 2023 facility, echo=FALSE, warning=FALSE}
ghgp2023 <- ghgp2023 %>%
  mutate(Emission_Range = cut(
    `Total.reported.direct.emissions`,
    breaks = c(0, 25000, 50000, 100000, 250000, 1000000, Inf),
    labels = c("0-25", "25-50", "50-100", "100-250", "250-1,000", ">1,000")
  ))

sector_summary <- ghgp2023 %>%
  separate_rows(`Industry.Type..sectors.`, sep = ",") %>%  # Split sectors
  mutate(`Industry.Type..sectors.` = trimws(`Industry.Type..sectors.`)) %>%  # Remove extra spaces
  group_by(Emission_Range, `Industry.Type..sectors.`) %>%
  summarise(
    Total_Emissions = sum(`Total.reported.direct.emissions`, na.rm = TRUE),
    .groups = "drop"
  )

range_counts <- ghgp2023 %>%
  group_by(Emission_Range) %>%
  summarise(Count = n(), .groups = "drop")

final_data <- sector_summary %>%
  left_join(range_counts, by = "Emission_Range")
```

```{r 2023 facility plot, echo=FALSE, warning=FALS}
# Plot
ggplot(final_data, aes(x = Emission_Range, y = Total_Emissions, fill = Industry.Type..sectors.)) +
  geom_bar(stat = "identity") +  # Stacked bar chart
  geom_text(data = range_counts, aes(x = Emission_Range, y = max(final_data$Total_Emissions) * 1.05, label = Count),
            inherit.aes = FALSE, size = 4) +  # Add count labels
  labs(
    title = "Facility Emission Ranges (2023)",
    x = "Emission Range of Facilities (thousand metric tons CO2e)",
    y = "2023 Emissions (million metric tons CO2e)",
    fill = "Sector"
  ) +
  theme_minimal() +
  custom_theme
```


### Fuel Type Emission Analysis [delete later -7 in ideas]





\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
