\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
setwd("~/EDE_FinalProject/Raw") 

# Load your packages
library(ggplot2)
library(dplyr)
library(knitr)
library(tidyr)
library(gt)
library(kableExtra)

# Set your ggplot theme
custom_theme <- theme(
  # Background theme customization
  panel.background = element_rect(fill = "white", color = "black"),  
  # Panel background color
  plot.background = element_rect(fill = "#f0f0f0", color = NA),  
  # Plot area background
  panel.grid.major = element_line(color = "gray", size = 0.5),  
  # Major grid lines
  panel.grid.minor = element_line(color = "lightgray", size = 0.25),  
  # Minor grid lines
  axis.line = element_line(color = "black"),  
  # Axis lines
  axis.text = element_text(size = 8, color = "black"),  
  # Axis tick labels
  axis.title = element_text(size = 10, color = "black"),  
  # Axis titles
  plot.title = element_text(size = 12, face = "bold", hjust = 0.5),  
  # Plot title customization
  legend.position = "none",  
  # Hide legend
  strip.background = element_rect(fill = "#D3D3D3", color = "black"),  
  # Strip background for facets
  strip.text = element_text(size = 10, face = "bold")  
  # Facet label text
)



# Load your datasets
fuel <- read.csv("Fuel_data.csv")
ghgp2010 <- read.csv("ghgp_data_2010.csv")
ghgp2011 <- read.csv("ghgp_data_2011.csv")
ghgp2012 <- read.csv("ghgp_data_2012.csv")
ghgp2013 <- read.csv("ghgp_data_2013.csv")
ghgp2014 <- read.csv("ghgp_data_2014.csv")
ghgp2015 <- read.csv("ghgp_data_2015.csv")
ghgp2016 <- read.csv("ghgp_data_2016.csv")
ghgp2017 <- read.csv("ghgp_data_2017.csv")
ghgp2018 <- read.csv("ghgp_data_2018.csv")
ghgp2019 <- read.csv("ghgp_data_2019.csv")
ghgp2020 <- read.csv("ghgp_data_2020.csv")
ghgp2021 <- read.csv("ghgp_data_2021.csv")
ghgp2022 <- read.csv("ghgp_data_2022.csv")
ghgp2023 <- read.csv("ghgp_data_2023.csv")
ghgpyear <- read.csv("ghgp_data_by_year_2023.csv")
```

# Rationale and Research Questions
>With current global temperature rise accelerating towards dangerously high levels, there is a significant threat to meeting the targets established by the Paris Climate Accords. A main target for this agreement was to limit global temperature to a 2.0° increase. According to the International Energy Agency, the United States and China together accounted for 45% of global fuel combustion emissions. (IEA). With the US being such a large contributor to climate change, a major goal of climate scientists is to find ways to reduce US GHG emissions. For this analysis, we chose to look at EPA-provided emission data from 2010-2023. The main purpose of this analysis is to gain insight on which industries demonstrate the greatest potential for significant emissions reductions, and to better understand what role do facility and fuel characteristics play in driving this change over the past decade?

>Industry sector data, greenhouse gas emission data, rated heat capacity data, and geospatial data will all be used to identify relationships and areas for improvement in the US. Additionally, most of our research questions are based on the most recent data, from 2023; however, time series analysis takes into account the full range of our data. Our central research questions set to examine how certain industry sectors should be limited in their production to limit the levels of greenhouse gas (especially carbon dioxide) emissions.

\newpage

# Dataset Information
>The dataset utilized in this analysis is sourced from the Greenhouse Gas Reporting Program (GHGRP), a program administered by the Environmental Protection Agency (EPA) under congressional mandate.
The GHGRP is the only comprehensive dataset that provides facility-level greenhouse gas (GHG) emissions data from major industrial sources across the United States. With over a decade of reporting (2010–2023) for most sectors, it offers critical insights into industrial emissions, including variations across industries, geographic regions, and time, down to the sector and facility levels.
The dataset includes detailed variables necessary for analyzing emissions trends and decarbonization progress across sectors. Key data points include:
Facility Information: Facility type, location, and rated heat capacity.
Emissions Data: Greenhouse gas type (e.g., CO₂, CH₄, N₂O), emission quantities, and fuel type.
Sector-Specific Details: Sector and subcategories (e.g., electricity generation, cement production).
The dataset also provides temporal data that enables longitudinal analysis, revealing trends in emissions reductions and variability across geographic regions and industry types.
The table below summarizes the structure of the dataset, including key variables, units, and their ranges or central tendencies:

```{r dataset info, echo=FALSE, warning=FALSE}
# Create the dataset structure as a data frame
data_info <- data.frame(
  Variable = c("Facility Type", "Location", "Rated Heat Capacity", 
               "GHG Type", "Emissions", "Fuel Type", "Year"),
  Description = c("Type of facility (e.g., power plant, refinery)",
                  "Facility location (state and region)",
                  "Facility’s energy production capacity",
                  "Type of greenhouse gas (e.g., CO₂, CH₄, N₂O)",
                  "Total emissions by facility",
                  "Type of fuel used (e.g., coal, natural gas, biomass)",
                  "Reporting year"),
  Unit = c("Categorical", "Categorical", "Million BTUs", 
           "Categorical", "Metric tons CO₂e", "Categorical", "Year")
)

# Create the table using kable
kable(data_info, 
      caption = "Table 1: Summary of Dataset Structure",
      col.names = c("Variable", "Description", "Unit" )) %>%
  kable_styling(full_width = FALSE, position = "center", bootstrap_options = c("striped", "hover"))

```


# Data Cleaning
>We used two main datasets, as well as a couple of shapefiles of the US. For the downloaded datasets, the main issue we encountered was that some facilities included multiple industry sector types. This made it difficult to group our data, so we decided to take the most significant sector (the first sector listed for each facility) and base our findings from that. Missing data values were excluded, and year/date format was checked to ensure the time series analysis would run smoothly. For the shapefile, we excluded all areas East of 50°West longitude, as the focus was on the US. After these initial steps were taken, we had confidence in the organization of our data and were ready for analysis.

>Downloaded datasets included date/time and mean discharge (cfs) columns. After the date column was properly classified as a date, unnecessary columns were removed from the data frame. Next, the date column was separated out to filter for years 1990-2021 and to filter for hurricane months (June-October). The dataset was then saved into the processed folder for each of the research locations. Missing data values were then filled using a linear interpolation and then daily averages of discharge were created in a new, clean column of daily means. After all fo these steps were taken the data was considered to be wrangled and was then ready for initial visualizations.

\newpage

# Exploratory Analysis 
> Our analysis focuses on assessing the performance of various sectors in decarbonization efforts, particularly their progress in reducing greenhouse gas (GHG) emissions. To facilitate this analysis, it is essential to first clarify the sector classifications within the dataset. Table 1 provides an overview of the Greenhouse Gas Reporting Program (GHGRP) sector classifications, grouping industrial activities into distinct categories such as Power Plants, Refineries, Chemicals, Fluorinated Chemicals, Waste, Metals, Minerals, Pulp & Paper, and Petroleum & Natural Gas Systems. Each sector is further broken down into specific subcategories, detailing the types of facilities and processes contributing to emissions, such as electricity generation under Power Plants or cement production under Minerals. This classification provides the foundation for understanding emission trends and evaluating decarbonization progress across sectors.

```{r sector information, echo=FALSE, warning=FALSE}
# Create a dataframe with the structure of the table
ghgrp_data <- data.frame(
  Sector = c("Power Plants", "Refineries", "Chemicals", "Fluorinated Chemicals", "Waste", 
             "Metals", "Minerals", "Pulp & Paper", "Petroleum & Natural Gas Systems – Direct Emissions"),
  Subcategories = c(
    "Electricity Generation",
    "Petroleum Refineries",
    "Adipic Acid Production; Ammonia Manufacturing; Hydrogen Production; Nitric Acid Production; Phosphoric Acid Production; Petrochemical Production; Silicon Carbide Production; Titanium Dioxide Production; Other Chemicals Production",
    "Fluorinated Gas Production; HCFC-22 Production/HFC-23 Destruction",
    "Municipal Landfills; Industrial Waste Landfills; Industrial Wastewater Treatment; Solid Waste Combustion",
    "Aluminum Production; Ferroalloy Production; Iron & Steel Production; Lead Production; Zinc Production; Magnesium Production; Other Metals Production",
    "Cement Production; Glass Production; Lime Manufacturing; Soda Ash Manufacturing; Other Minerals Production",
    "Chemical Pulp & Paper; Other Paper Producers",
    "Onshore Production; Offshore Production; Gathering and Boosting; Natural Gas Processing; Natural Gas Transmission Compression; Natural Gas Transmission Pipelines; Natural Gas Distribution; Underground Natural Gas Storage; Liquefied Natural Gas Storage; Liquefied Natural Gas Import/Export; Other Petroleum and Natural Gas Systems"
  )
)

# Split the long list into separate rows for better readability (if needed)
ghgrp_data <- ghgrp_data %>%
  mutate(Subcategories = strsplit(as.character(Subcategories), "; ")) %>%
  tidyr::unnest(Subcategories)

# Create the styled table using the `gt` package
ghgrp_table <- gt(ghgrp_data) %>%
  tab_header(
    title = "GHGRP Sector Classifications"
  ) %>%
  cols_label(
    Sector = "Sector",
    Subcategories = "Subcategories"
  ) %>%
  tab_options(
    table.width = px(800),
    heading.title.font.size = px(18),
    column_labels.font.weight = "bold"
  )

# Create the table and add a caption
kable(
  ghgrp_data,
  caption = "Table 2: Classification of sectors under the Greenhouse Gas Reporting Program (GHGRP)."
) %>%
  kable_styling(full_width = FALSE, position = "left")
```


>The table 3 provides an overview of the distribution of reporting facilities across various industry sectors in the 2023 Greenhouse Gas Reporting Program (GHGRP) dataset. To create this summary, the industry sectors associated with each facility were split into individual categories and grouped for analysis. The number of reporters in each sector was then calculated, highlighting the contribution of different industries to overall emissions reporting. This breakdown offers valuable insights into the sectors with the most active reporting and provides a foundation for further sector-specific analysis of emissions trends and decarbonization progress.

```{r 2023 sector summary, echo=FALSE, warning=FALSE, message=FALSE}
# Split the "Industry.Type..sectors." column by commas and unnest into separate rows
industry_summary <- ghgp2023 %>%
  separate_rows(`Industry.Type..sectors.`, sep = ",") %>%  # Correct column name
  mutate(`Industry.Type..sectors.` = trimws(`Industry.Type..sectors.`)) %>%  # Remove extra spaces
  group_by(`Industry.Type..sectors.`) %>%  # Group by the individual sectors
  summarise(Number_of_Reports = n())  # Count the number of reports for each sector

# Display the summarized data as a styled table
kable(industry_summary, 
      caption = "Table 3: Summary of Number of Reporters by Industry Sector", 
      col.names = c("Industry Sector", "Number of Reports")) %>%
  kable_styling(full_width = FALSE, position = "center", bootstrap_options = c("striped", "hover"))
```

>The bar chart (figure 1) illustrates the distribution of emissions by gas type. The chart highlights the dominant contributions of carbon dioxide (CO₂), with methane (CH₄) and nitrous oxide (N₂O) playing significant roles as well. This visualization emphasizes the need for sector-specific strategies to address each type of GHG.

```{r 2023 GHG summary, echo=FALSE, warning=FALSE, message=FALSE}
# Reshape the data into long format
ghgp_long <- ghgp2023 %>%
  select(
    CO2.emissions..non.biogenic., 
    Methane..CH4..emissions, 
    Nitrous.Oxide..N2O..emissions, 
    HFC.emissions, 
    PFC.emissions, 
    SF6.emissions, 
    NF3.emissions, 
    Other.Fully.Fluorinated.GHG.emissions
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = "Gas_Type",
    values_to = "Emissions"
  )
```

```{r 2023 GHG plot, echo=FALSE, warning=FALSE}
# Create the bar chart with the custom theme
ggplot(ghgp_long, aes(x = Emissions, y = Gas_Type, fill = Gas_Type)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Greenhouse Gas Emissions in 2023",
    x = "Greenhouse Gas Type",
    y = "Emissions (Tons of CO2)",
    caption = "Figure 1: Distribution of greenhouse gas emissions by gas type in 2023."
  ) +
  custom_theme +  # Apply the custom theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  )
```

\newpage

# Analysis


## Question 1: Which industries demonstrate the greatest change for significant emissions reductions over the past decade? 


### 2023 Average CO2 Emissions by Industry Sector 
>Daniel:

```{r 2023 Industry Sector plot, echo=FALSE, warning=FALSE, message=FALSE}
# Calculate average CO2 emissions by industry sector
 CO2_2023 <- ghgp2023 %>%
  separate_rows(`Industry.Type..sectors.`, sep = ",") %>%  # Correct column name
  mutate(`Industry.Type..sectors.` = trimws(`Industry.Type..sectors.`)) %>%  # Remove extra spaces
  group_by(`Industry.Type..sectors.`) %>%  # Group by the individual sectors
  summarise(Average_CO2 = mean(CO2.emissions..non.biogenic., na.rm = TRUE) / 10^6)

CO2_2023$Industry.Type..sectors. <- gsub("Fluorintaed", "Fluorinated", CO2_2023$Industry.Type..sectors.)
CO2_2023$Industry.Type..sectors. <- gsub("Import and Export", "Distribution", CO2_2023$Industry.Type..sectors.)
# Generate the bar chart
ggplot(CO2_2023, aes(x = reorder(Industry.Type..sectors., -Average_CO2), y = Average_CO2, fill = Industry.Type..sectors.)) +
  geom_bar(stat = "identity") +
  labs(title = "Average 2023 CO2 Emissions by Industry Sector",
       x = "Industry Sector",
       y = "Average CO2 Emissions (MMT)",
       caption= "Figure 2: Average 2023 CO2 Emissions by Industry Sector"
       ) +
  custom_theme +  # Apply the custom theme
  theme(
    axis.text.x = element_text(hjust = 1)  # Rotate x-axis labels
  )+
 coord_flip() 
```
### 2010-2023 Average CO2 Emissions by Industry Sector 
>The emission trend 2010-2023 is shown in figure 3. Most industry stays stable but the refinery goes higher and higher. For sources reporting to the GHGRP, emissions decreased 4.4% from 2022 to 2023. Between 2011 and 2023, GHGRP-reported direct emissions (e.g. excluding suppliers) decreased 22.3%. This decline is primarily caused by the decline in reported emissions from power plants, which decreased 33.8% over the same period.


```{r time series, echo=FALSE, warning=FALSE, message=FALSE}
emissions_data <- ghgpyear %>%
  select(Latest.Reported.Industry.Type..sectors., 
         X2023.Total.reported.direct.emissions,
         X2022.Total.reported.direct.emissions,
         X2021.Total.reported.direct.emissions,
         X2020.Total.reported.direct.emissions,
         X2019.Total.reported.direct.emissions,
         X2018.Total.reported.direct.emissions,
         X2017.Total.reported.direct.emissions,
         X2016.Total.reported.direct.emissions,
         X2015.Total.reported.direct.emissions,
         X2014.Total.reported.direct.emissions,
         X2013.Total.reported.direct.emissions,
         X2012.Total.reported.direct.emissions,
         X2011.Total.reported.direct.emissions) %>%
  pivot_longer(cols = starts_with("X"), names_to = "Year", values_to = "Emissions") %>%
  group_by(Year, Latest.Reported.Industry.Type..sectors.) %>%
  summarize(Emissions = sum(Emissions, na.rm = TRUE), .groups = "drop")

# Clean the year column
emissions_data$Year <- as.numeric(gsub("X(\\d{4}).*", "\\1", emissions_data$Year))

# Plot the graph
ggplot(emissions_data, aes(x = Year, y = Emissions / 1e6, color = Latest.Reported.Industry.Type..sectors.)) +
  geom_line(size = 1) +
  labs(
    title = "2010-2023 Average CO2 Emissions by Industry Sector",
    x = "Year",
    y = "Emissions (million metric tons CO2e)",
    caption="Figure 3: 2010-2023 Average CO2 Emissions by Industry Sector",
    color = "Sector"
  ) +
  custom_theme 
```



## Question 2: spatial analysis
>Daniel:

### 2023  Total CO2 Emissions by Industry Sector and State

```{r 2023 Industry Sector by State plot, echo=FALSE, warning=FALSE, message=FALSE}

CO2_by_state_sector <- ghgp2023 %>%
  separate_rows(`Industry.Type..sectors.`, sep = ",") %>%
  mutate(`Industry.Type..sectors.` = trimws(`Industry.Type..sectors.`)) %>%
  group_by(State, `Industry.Type..sectors.`) %>% 
  summarise(Total_CO2 = sum(CO2.emissions..non.biogenic., na.rm = TRUE) / 10^6) 
  
CO2_by_state_sector$Industry.Type..sectors. <- gsub("Fluorintaed", "Fluorinated", CO2_by_state_sector$Industry.Type..sectors.)
CO2_by_state_sector$Industry.Type..sectors. <- gsub("Import and Export", "Distribution", CO2_by_state_sector$Industry.Type..sectors.)
# Generate the stacked bar chart
ggplot(CO2_by_state_sector, 
       aes(x = reorder(State, -Total_CO2), 
           y = Total_CO2, 
           fill = `Industry.Type..sectors.`)) +
  geom_bar(stat = "identity") + 
  labs(title = "Total 2023 CO2 Emissions by State and Industry Sector",
       x = "State",
       y = "Total CO2 Emissions (MMT)",
       caption = "Figure 4: Total 2023 CO2 Emissions by State and Industry Sector ") +
  custom_theme + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()

```

### 2023 CO2 Emissions by State

```{r 2023 Industry Sector plot, echo=FALSE, warning=FALSE, message=FALSE}

library(sf)
# Filter out rows with Longitude greater than 100
ghgp2023 <- ghgp2023[ghgp2023$Longitude <= 60, ]

ghgp2023 <- ghgp2023[!is.na(ghgp2023$Longitude) & !is.na(ghgp2023$Latitude), ]
USnation_sf<- st_read("Raw/cb_2018_us_nation_20m.shp")
st_crs(USnation_sf)

ghgp2023_sf <- ghgp2023 %>%
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4269) 

str(ghgp2023_sf)

st_crs(ghgp2023_sf)

ggplot() +
  geom_sf(data = USnation_sf) + 
  geom_sf(data = ghgp2023_sf,
          aes(size = CO2.emissions..non.biogenic.,
              color = CO2.emissions..non.biogenic.),  # Map CO2 emissions to color 
          alpha = 0.5) + 
  labs(title = "CO2 Emissions from Gas Plants",
       size = "CO2 Emissions",
       caption = "Figure 5: Map of CO2 emissions from gas plants across the United States"
       ) +
  coord_sf(xlim = c(-180, -50)) +  # Set longitude limits
  scale_color_gradient(low = "cadetblue1", high = "grey1")+  # Define the color scale
  scale_size(range = c(1,7)) 

```


#### 2023 Direct Emission Facilities by State
```{r 2023 Facilities plot, echo=FALSE, warning=FALSE, message=FALSE}
# Calculate average CO2 emissions by industry sector
units_data <- read.csv("Raw/ede_csv_unit_emissions.csv", skip = 6)
#RHC_by_state <- units_data %>% 
  #group_by(State) %>% 
  #summarise(Average_RHC = mean(Unit Maximum Rated Heat Input Capacity (mmBTU/hr), na.rm = TRUE)
            
Facilities_by_state <- units_data %>% 
  filter(`Reporting.Year` == 2023) %>% 
  group_by(State) %>% 
  summarise(FacilityCount = n_distinct(Facility.Id)) %>%
  arrange(desc(FacilityCount))

# Create the bar chart
ggplot(Facilities_by_state, aes(x = reorder(State, -FacilityCount), y = FacilityCount, fill = FacilityCount)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Number of Unique Facilities by State (2023)",
    x = "State",
    y = "Number of Facilities",
    caption = "figure 6: Number of Unique Facilities by State (2023)"
  )

```

## Question 3: What role do facility and fuel characteristics play in driving this change over the past decade?

### Facility Emission Analysis 
>The GHGRP provides a robust dataset that can be used to determine the number of facilities at various emissions
levels in many industry sectors. Figure 7 shows the emission ranges in 2023. The number on the bar shows the amount of facilities in that range. Most of reporting facilities had emissions less than 250,000 metric tons CO2e. In 2023, the 633 largest-emitting facilities—those emitting more than one million metric tons CO2e—accounted for approximately 1.71 billion metric tons of CO2e. These high-emitting facilities are mainly power plants, but they also include facilities in all other direct emitter sectors.

```{r 2023 facility, echo=FALSE, warning=FALSE}
ghgp2023 <- ghgp2023 %>%
  mutate(Emission_Range = cut(
    `Total.reported.direct.emissions`,
    breaks = c(0, 25000, 50000, 100000, 250000, 1000000, Inf),
    labels = c("0-25", "25-50", "50-100", "100-250", "250-1,000", ">1,000")
  ))

sector_summary <- ghgp2023 %>%
  separate_rows(`Industry.Type..sectors.`, sep = ",") %>%  # Split sectors
  mutate(`Industry.Type..sectors.` = trimws(`Industry.Type..sectors.`)) %>%  # Remove extra spaces
  group_by(Emission_Range, `Industry.Type..sectors.`) %>%
  summarise(
    Total_Emissions = sum(`Total.reported.direct.emissions`, na.rm = TRUE),
    .groups = "drop"
  )

range_counts <- ghgp2023 %>%
  group_by(Emission_Range) %>%
  summarise(Count = n(), .groups = "drop")

final_data <- sector_summary %>%
  left_join(range_counts, by = "Emission_Range")
```

```{r 2023 facility plot, echo=FALSE, warning=FALS}
# Plot
ggplot(final_data, aes(x = Emission_Range, y = Total_Emissions, fill = Industry.Type..sectors.)) +
  geom_bar(stat = "identity") +  # Stacked bar chart
  geom_text(data = range_counts, aes(x = Emission_Range, y = max(final_data$Total_Emissions) * 1.05, label = Count),
            inherit.aes = FALSE, size = 4) +  # Add count labels
  labs(
    title = "Facility Emission Ranges (2023)",
    x = "Emission Range of Facilities (thousand metric tons CO2e)",
    y = "2023 Emissions (million metric tons CO2e)",
    caption = "Figure 7: Facility Emission Ranges (2023)",
    fill = "Sector"
  ) +
  theme_minimal() +
  custom_theme
```


### Fuel Type Emission Analysis
>Figure 8 illustrates the emissions of methane (CH4) and nitrous oxide (N2O) across various fuel types, measured in metric tons of CO2 equivalent (mt CO2e). Coal exhibits the highest emissions for both CH4 and N2O, with methane emissions being notably dominant. Other fuel types, such as petroleum products and solid byproducts, have significantly lower emissions in comparison, while natural gas shows relatively minimal emissions in both categories. 

```{r fuel type, echo=FALSE, warning=FALSE}
fuel_data <- fuel %>%
  select(General.Fuel.Type, 
         Fuel.Methane..CH4..emissions..mt.CO2e., 
         Fuel.Nitrous.Oxide..N2O..emissions..mt.CO2e.) %>%
  group_by(General.Fuel.Type) %>%
  summarize(
    CH4_Emissions = sum(Fuel.Methane..CH4..emissions..mt.CO2e., na.rm = TRUE),
    N2O_Emissions = sum(Fuel.Nitrous.Oxide..N2O..emissions..mt.CO2e., na.rm = TRUE)
  ) %>%
  pivot_longer(cols = starts_with("CH4_Emissions"):starts_with("N2O_Emissions"), 
               names_to = "Emission_Type", 
               values_to = "Emissions") 

# Plot the grouped bar chart
ggplot(fuel_data, aes(x = General.Fuel.Type, y = Emissions, fill = Emission_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "CH4 and N2O Emissions by Fuel Type",
    x = "Fuel Type",
    y = "Emissions (mt CO2e)",
    caption = "Figure 8: GHG Emissions by Fuel Type in 2023",
    fill = "Emission Type"
  ) +
  custom_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```




## Question 4: Rated heat capacity
>Daniel:

### 2023 Average Rated Heat Capacity by State
```{r 2023 AVG Rated Heat Capacity plot, echo=FALSE, warning=FALSE, message=FALSE}
# Calculate RHC by state

AVG_RHC_by_state <- units_data %>% 
  filter(`Reporting.Year` == 2023) %>% 
  group_by(State) %>% 
  summarise(Average_RHC = mean(Unit.Maximum.Rated.Heat.Input.Capacity..mmBTU.hr., na.rm = TRUE))%>%
  arrange(desc(Average_RHC))


# Create the bar chart
ggplot(AVG_RHC_by_state, aes(x = reorder(State, -Average_RHC), y = Average_RHC, fill = Average_RHC)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Average Rated Heat Capacity by State (2023)",
    x = "State",
    y = "Rated Heat Capacity",
    caption = "Figure 9:Average Rated Heat Capacity by State (2023) "
  )
```

### 2023 Total Rated Heat Capacity by State
```{r 2023 Total Rated Heat Capacity plot, echo=FALSE, warning=FALSE, message=FALSE}
Total_RHC_by_state <- units_data %>% 
  filter(`Reporting.Year` == 2023) %>% 
  group_by(State) %>% 
  summarise(Total_RHC = sum(Unit.Maximum.Rated.Heat.Input.Capacity..mmBTU.hr., na.rm = TRUE)) %>%
  arrange(desc(Total_RHC))
# Create the bar chart
ggplot(Total_RHC_by_state, aes(x = reorder(State, -Total_RHC), y = Total_RHC, fill = Total_RHC)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Total Rated Heat Capacity by State (2023)",
    x = "State",
    y = "Rated Heat Capacity (MMBTU/hr)",
    caption = "Figure 10:Total Rated Heat Capacity by State (2023) "
  )
```





\newpage

# Summary and Conclusions
> Question 1: Which industries demonstrate the greatest change for significant emissions reductions over the past decade?

> The analysis of greenhouse gas emissions from 2010 to 2023 reveals significant trends and insights into the sources and drivers of emissions. While most industry sectors maintained stable emissions over this period, the refinery sector experienced a consistent increase. Overall, GHGRP-reported direct emissions decreased by 22.3% between 2011 and 2023, with a notable 4.4% decline from 2022 to 2023. This reduction is largely driven by a 33.8% decline in emissions from power plants, which remain the largest contributors among high-emitting facilities.



>Question 3: What role do facility and fuel characteristics play in driving this change over the past decade?

>Facility-level data highlights the disproportionate impact of a small subset of facilities. In 2023, the 633 largest emitters, primarily power plants, accounted for 1.71 billion metric tons of CO2e, underscoring the importance of targeting these high-emitting sources in mitigation strategies.
Fuel type analysis further illustrates coal's significant role in greenhouse gas emissions, particularly for methane (CH4) and nitrous oxide (N2O). These findings emphasize the need for tailored approaches that address both facility and fuel characteristics. Transitioning to cleaner fuels like natural gas and reducing emissions from coal-intensive industries will be critical in driving further reductions in greenhouse gas emissions over the next decade.

\newpage

# References
[1] “Greenhouse Gas Emissions from Energy Data Explorer – Data Tools.” IEA, 2 Aug. 2024, www.iea.org/data-and-statistics/data-tools/greenhouse-gas-emissions-from-energy-data-explorer. 
[2] “GHGRP 2023 Overview Profile.” EPA, October 2024, www.epa.gov/system/files/documents/2024-10/ghgrp-2023-overview-profile.pdf.
